Thesis: draft

### 2021-05-21 draft.tex
Update draft.tex
changed beginning of results stubborn actors

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_1000-6_PR_20-80_8x7.png}
    \label{ws1000}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10000-6_PR_20-80_8x7.png}
    \label{ws10000}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for unbalanced starting conditionsn for a WS model with $N=10^3$ or $N=10^4$, $\left<k\right>=6$ and $\beta = 0.01$. The PR filtering algorithm is used.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for unbalanced starting conditions. The Watts-Strogatz network with $\left<k\right> = 6$ and $\beta = 0.01$ is depicted with \textbf{(a)} $N=10^3$; \textbf{(b)} $N=10^4$. In both networks the average clustering coefficient is $\left<cc\right> = 0.584 \pm 0.003$. The PR filtering filtering algorithm is used. The dotted black represents the initial prevalence of opinion A. The dotted grey line represents a vanished prevalence of opinion A. Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization. Bars represent the standard deviation $\sigma$.}
\label{ev_op_WS_k=6}
\end{figure}
\noindent
\textbf{Echo chambers}\\
\newline
Fig. \ref{echo_20_80} shows the distribution of the fraction of friends of actor $i$ with the same opinion B at $t=500$.  The REC and PR filtering algorithms are compared for the ER, WS, SBM and SBM-WS.\\
\newline
The results for the ER model and the SBM (both the high and low modularity configuration) look qualitatively the same. Similarly, the results for the WS model and the SBM-WS agree with each other as well. This difference in results between ER/SBM (negligible clustering) and WS/SBM-WS (high clustering) suggests that it is the clustering that drives these differences.

@@

  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t=500$, $\left<P_\text{B}^{\text{nn}}\right>$, for unbalanced starting conditions.]{Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$ for unbalanced starting conditions. \textbf{(a)} ER; \textbf{(b)} WS; \textbf{(c)} SBM; \textbf{(d)} SBM-WS. Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$). For each model the REC and the PR filtering algorithms are compared. The dotted black line represents a constant amount of echo chambers over time. The distribution is normalized by dividing for the same quantity computed at $t=0$. Each curve is the average of $10^2$ independent simulations. Error bars are omitted for improved visualization.}

@@

In this section, we will study the effect of stubborn actors on the opinion dynamics. We focus on two different stubbornness models, both are thoroughly discussed in Chapter \ref{chap3}, Section \ref{stubbBin}. The results of the first model, including a stubbornness parameter, are given in Section \ref{stubbpar}. The results of the second model, the majority threshold model, are presented in Section \ref{majThres}.\\
\newline
We will implement four different network models in order to investigate the possible interplay between network structure and stubbornness. Each of the network models have a total number of nodes $N=10^3$ and average degree $\left<k\right> \sim 10$. The baseline model is the Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$. The interplay between stubbornness and clustering is studied with the Watts-Strogatz model (WS) with mean degree $\left<k\right>=10$, rewire probability $\beta = 0.06$ and clustering coefficient $\left<cc\right> = 0.557 \pm 0.007$. The effect of community structure is investigated with the stochastic block model (SBM) with 10 communities of size 100. Two different modularities are implemented: a high modularity, $Q = 0.908 \pm 0.005$, with $p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$) and a low modularity, $Q = 0.220 \pm 0.007$, with $p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$. Finally, the interplay between stubbornness and both clustering and community structure is studied with the stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100. Each community has the following Watts-Strogatz parameters: mean degree $\left<k\right> = 10$ and rewire probability $\beta = 0.01$. Edges are added between communities with a probability $p_{\text{add}} = 0.001$. The clustering coefficient of this network is $\left<cc\right> = 0.553 \pm 0.004$ and the modularity is $Q = 0.909 \pm 0.004$.\\
\newline
For each network model, we compare the REF, REC and PR filtering algorithms to analyze the possible interplay between stubbornness and filtering algorithm.

\subsection{Including a stubbornness parameter}
\label{stubbpar}

In this section, we study the effect of a stubbornness parameter $r_i$ on two configurations: (i) fraction of actors completely stubborn, (ii) all actors partly stubborn.\\
\newline
In both cases we study both the evolution of the prevalence of the opinions and the formation of echo chambers. This is done for two different starting conditions: balanced starting conditions and unbalanced starting conditions. \\
\newline
The opinion dynamics model described in Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented with an activation probability $p_{\text{act}} = 0.1$ and a maximum length of the actual time-line $S = 20$.\\
\newline
\textbf{Completely stubborn actors}\\
\newline
We make a fraction of the actors (fracRes) completely stubborn ($r_i = 1$). We go from no stubborn actors, fracRes $=0$, to all the actors are stubborn, fracRes $=1$. In all cases, the stubborn actors are randomly distributed over the network.\\
\newline
We will first discuss the formation of echo chambers for an initial $50/50$ opinion distribution, after which the results for the opinion evolution will be shared. Then we will discuss the formation of echo chambers and the evolution of the prevalence of the opinions for an initial $20/80$ opinion distribution.

@@

	\caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR filtering algorithm and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $50/50$ opinion distribution and using the PR filtering algorithm. The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001$, $Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008$, $Q = 0.220 \pm 0.007$) are depicted; ER with $p= 0.01$. Each data point is the average of $10^2$ independent simulations. Bars represent the standard error $SE$.}

>>>>

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_1000-6_PR_20-80_8x7.png}
    \label{ws1000}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10000-6_PR_20-80_8x7.png}
    \label{ws10000}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for unbalanced starting conditions. A WS model with $N=10^3$ or $N=10^4$, $\left<k\right>=6$ and $\beta = 0.01$ is implemented. The PR filtering algorithm is used.]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for unbalanced starting conditions. The Watts-Strogatz network with $\left<k\right> = 6$ and $\beta = 0.01$ is depicted with \textbf{(a)} $N=10^3$; \textbf{(b)} $N=10^4$. In both networks the average clustering coefficient is $\left<cc\right> = 0.584 \pm 0.003$. The PR filtering filtering algorithm is used. The dotted black represents the initial prevalence of opinion A. The dotted grey line represents a vanished prevalence of opinion A. Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization. Bars represent the standard deviation $\sigma$.}
\label{ev_op_WS_k=6}
\end{figure}
\noindent
\textbf{Echo chambers}\\
\newline
Fig. \ref{echo_20_80} shows the distribution of the fraction of friends of actor $i$ with the same opinion B at $t=500$.  The REC and PR filtering algorithms are compared for the ER, WS, SBM and SBM-WS.\\
\newline
The results for the ER model and the SBM (both the high and low modularity configuration) look qualitatively the same. Similarly, the results for the WS model and the SBM-WS agree with each other as well. This difference in results between ER/SBM (negligible clustering) and WS/SBM-WS (high clustering) suggests that it is the clustering that drives these differences.

@@

  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t=500$, $\left<P_\text{B}^{\text{nn}}\right>$, for unbalanced starting conditions.]{Distribution of the fraction of friends (nearest neighbors, nn) of actor $i$ with the same opinion B at $t = 500$ for unbalanced starting conditions. \textbf{(a)} ER; \textbf{(b)} WS; \textbf{(c)} SBM; \textbf{(d)} SBM-WS. The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$). Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$). For each model the REC and the PR filtering algorithms are compared. The dotted black line represents a constant amount of echo chambers over time. The distribution is normalized by dividing for the same quantity computed at $t=0$. Each curve is the average of $10^2$ independent simulations. Error bars are omitted for improved visualization.}

@@

Stubbornness is an important feature of real-life opinion dynamics. Understanding the impact of individual resistance and the interplay with filtering algorithms and network topologies is crucial. In this section, we will study the effect of stubborn actors on the opinion dynamics. We focus on two different stubbornness models, discussed in Section \ref{stubbBin}. The results of the first model, including a resistance parameter, are given in Section \ref{stubbpar}. The results of the second model, the majority threshold model, are presented in Section \ref{majThres}.\\
\newline
We will compare the REF, REC and PR filtering algorithms to analyze the possible interplay between stubbornness and filtering algorithm. The ER, WS, SBM and SBM-WS are implemented to study the interplay between network structure and stubbornness.

\subsection{Resistance parameter}
\label{stubbpar}

We study the effect of a stubbornness parameter $r_i$ on two configurations: (i) fraction of actors completely stubborn, (ii) all actors partly stubborn. In both cases we study both the evolution of the prevalence of the opinions and the formation of echo chambers. This is done for balanced and unbalanced starting conditions. \\
\newline
\textbf{Completely stubborn actors}\\
\newline
We make a fraction of the actors ($f_s$) completely stubborn ($r_i = 1$). We go from no stubborn actors, fracRes $=0$, to all the actors are stubborn, fracRes $=1$. \\
\newline
We will first discuss the formation of echo chambers for balanced starting conditions, after which the results for the opinion evolution will be shared. Then we will discuss the formation of echo chambers and the evolution of the prevalence of the opinions for unbalanced starting conditions.

@@

	\caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR filtering algorithm and balanced starting conditions.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for balanced starting conditions and using the PR filtering algorithm. Results are shown for: (i) WS; (ii) SBM-WS; (iii) SBM and (iv) ER. The WS model has $\beta = 0.06$ ($\left<cc\right> = 0.557 \pm 0.007$). Two configurations are implemented for the SBM: a high modularity configuration ($Q = 0.908 \pm 0.005$)) and a low modularity configuration ($Q = 0.220 \pm 0.007$). The $y$-axis shows the average of the normalized fraction of actors with all neighbors having opinion A and the normalized fraction of actors with all neighbors having opinion B. Each data point is the average of $10^2$ independent simulations. Bars represent the standard error $SE$.}
